---
title: "p8105_hw3_qz2493"
author: "Qingyue Zhuo qz2493"
date: "2022-10-11"
output: github_document
---
# Load packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

library(p8105.datasets)
data("ny_noaa")

library(ggridges)
```

# Problem 1
## Load the instacard data
```{r}
library(p8105.datasets)
data("instacart")
```


# Problem 2
## Load and tidy the "accel" dataset
```{r}
accel_data = 
  read_csv("data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity",
    values_to = "observations") %>%
  mutate(
    day = factor(day),
    day = fct_relevel(day, "Monday", "Tuesday", "Wednesday", "Thursday", "Friday","Saturday", "Sunday"),
    day_type = case_when(
      day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "Weekday",
      day %in% c("Saturday", "Sunday") ~ "Weekend"
    )) %>%
  arrange(week, day)
      
```
### This "accel" dataset has `r nrow(accel_data)` observations of `r ncol(accel_data)` variables. The names of the variables are `r names(accel_data)`.

## Total activiteis over the day
```{r}
accel_data %>%
  group_by(week, day) %>%
  summarize(total_activity = sum(observations)) %>%
  pivot_wider(
    names_from = "day",
    values_from = "total_activity"
  ) %>%
  knitr::kable()
```

## 24-hour-activity
```{r}
accel_data %>%
  mutate(
    activity = substring(activity, 10),
    activity = as.integer(activity)
  ) %>%
  ggplot(aes(x = activity, y = observations, color = day)) +
  geom_line(alpha = 0.5) +
  scale_x_continuous(
    breaks = c(0, 120, 240, 360, 480, 600, 720, 840, 960, 1080, 1200, 1320, 1440),
    labels = c("0", "2", "4", "6", "8", "10", "12", "14", "16", "18", "20", "22", "24")) +
  labs(
    title = "24-Hour Activity for Each Day",
    x = "Time of Day",
    y = "Activity Level"
  ) +
  theme(legend.position = "bottom")
  
```

# Problem 3
## Load and tidy the noaa data
```{r}
ny_data = 
  ny_noaa %>%
  janitor::clean_names() %>%
  separate(date, into = c("year", "month", "day"), sep = "-")
```

```{r}
ny_data =
  ny_data %>%
  rename(snow_mm = "snow",
         prcp_mm = "prcp",
         snwd_mm = "snwd",
         tmax_F = "tmax",
         tmin_F = "tmin")

```

```{r}
ny_data = 
  ny_data %>%
    mutate(tmin_F = as.numeric(tmin_F),
          tmax_F = as.numeric(tmax_F))
```

```{r}
df1 = data.frame(ny_data)
names(which.max(table(df1$snow_mm, useNA = "ifany"))) 
```
###

### how to ensure that observations for temperature, precipitation, and snowfall are given in reasonable units

### what are the most commonly observed values? why?
```{r}

```

### two- panel- plot-
     a. average max temperature in January in each station across the year
     b. average max temperature in July in each station across the year
```{r}

```

## two-panel-plot
    a. tmax vs. tmin for the full dataset- non-scatter plot
    b. distribution of snowfall values greater than 0 and less than 100 separately by year


```{r}
plot_1 = 
  ny_data %>%
  group_by(id, year, month) %>%
  filter(month %in% c("07","01")) %>%
  summarize(tmax_mean = mean(tmax_F,na.rm = TRUE, color = id)) %>%
  ggplot(aes(x = year, y = tmax_mean,group = id)) +
  geom_point() + geom_path() +
  facet_grid( ~ month)

plot_1
```

```{r}
plot2 = 
  ny_data %>%
  ggplot(aes(x = tmax_F, y = tmin_F)) +
  geom_hex()


plot3 = 
  ny_data %>%
  group_by(id, year, month) %>%
  filter( snow_mm < 100 & snow_mm > 0) %>%
  ggplot(aes(x = snow_mm, y = year)) +
  geom_density_ridges()


```



